A Full Accounting of Landcover Map Error and Bias and Their Impacts on Assessments of Global Change

### 2015-10-10 paper/main.tex
October 10th, 2015 commit. Edits in changelog.txt.

%\usepackage{pnastwoF}
%\usepackage{cite}
\usepackage[numbers,round,sort,compress]{natbib}
\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
\usepackage{lineno}
\setlength\linenumbersep{2pt}
\usepackage{placeins}
\usepackage{float}

@@

Being unable to fully quantify the errors in landcover maps of course makes it difficult, if not impossible, to quantify their impact on downstream analyses. There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, and carbon flux \cite{quaife_impact_2008} and human population estimates \cite{linard_assessing_2010}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}. One exception is a study  \cite{schmit_limitations_2006} that used a high quality, ground-collected reference map detailing farm land use parcels in central Belgium, but the number of sites and region were both fairly restricted, and the parcels were not spatially contiguous.

Just as a building needs a solid foundation, global change science needs to be based on sound landcover data. There is thus an urgent need to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.  We address this need in this study, using a unique, high accuracy agricultural landcover map for the entire country of South Africa to conduct a spatially comprehensive, bottom-up quantification of error in several latest generation landcover maps that are widely used in global change studies. We use these errors to assess the extent of bias in i) landcover data, ii) how  landcover properties influence this bias, iii) how these biases change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and iv) how these biases propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.

@@

The South African government commissioned a whole country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}. The map was made by trained workers who visually interpreted high resolution satellite imagery ($<$5 m SPOT imagery) and manually digitized field boundaries following a standardized mapping protocol. The resulting vectorized field maps, which were made in 2007 and updated in 2011, provide a unique, high quality reference dataset describing crop field distributions and size classes, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution. We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create a gridded cropland reference map.

We compared the reference map with similar maps derived from four existing landcover datasets. We obtained South Africa's 30 m resolution National Landcover map (SA-LC) for 2009 \cite{sanbi_national_2009}, the 500 m resolution MODIS Landcover for 2011 \cite{land_processes_distributed_active_archive_center_lp_daac_modis_2011, friedl_modis_2010}, the 300 m resolution GlobCover 2009 \cite{arino_global_2012}, and the new 1 km Geo-wiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}. We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite[e.g.][]{fry_completion_2009}, MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013, shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and represents the current state-of-the-art in landcover mapping. We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates (hereafter simply the ``cropland maps''), resulting in 4 maps to compare to our reference map.

\vspace{-0.5 cm}
\section{Quantifying Error and Bias}
We first quantified the errors in cropland area estimates based on the pixel-wise differences between the reference map and each of the four cropland maps. We calculated these errors for five different scales of aggregation, from the original 1 km up to 100 km, in order to calculate bias (the mean pixel-wise error) and mean absolute error (which measures the average error magnitude), and how these vary with scale. Next, we assessed the degree to which the average cropland cover in agricultural landscapes, a descriptor of landcover pattern, impacts error magnitude.

We then investigated how landcover map error propagates through ``downstream'' analyses that are built on landcover data, and which are typical of those performed in global change assessments. We began by examining the biases and mean absolute errors in three ``first-order'' landcover-based analyses, in which a variable of interest is mapped onto a landcover type(s) using a simple empirical relationship. The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}. The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates. Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.

Finally, we examined errors resulting from two second-order analyses, in which a process model draws on the cover types' values to calculate an output value. For the first of these, we used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model to calculate monthly evapotranspiration, using the reference and cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components. In the second example, we examined how these map errors impact the parameterization of an agent-based food security model \cite{chen_dependency_2013}. Spatially-explicit, agent-based models are frequently employed in land change science, and require an initialization step to assign landscape resources to model agents  \cite[e.g.][]{manson_agent-based_2007,evans_multi-scale_2004,kelley_relative_2011}. In this case, we used cropland maps to provide the model the location and abundance of cropland, which is used to  allocate an initial share of cropland to each farm household (agent) in the model.

\vspace{-0.5 cm}
\section{Bias}
\subsection{Landcover error and bias}

We created the 1 km reference map after removing all field types classified as communal/smallholder agriculture (individual fields in this category were not delineated, thus they were removed to prevent potential cropland area overestimates, SI) or permanent tree-crops (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa). The 2011 reference map showed a cropland area of 104,310 km$^2$, which the SA-LC and GeoWiki maps overestimated by 26 and 5.8\%, respectively, and GlobCover and MODIS underestimated by 21 and 26.1\%.

We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from each reference (2007 and 2011) map at each scale of aggregation to assess error patterns (Fig. 1). Negative pixels here represent overestimation error, while positive values indicate underestimates.

\vspace{-0.25 cm}
\begin{figure}[ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/bias_map.pdf}}
\caption{Errors in the percent cropland estimates resulting from each of the four cropland maps relative to the reference map at different scale of pixel aggegation. Rows indicate the landcover map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation. White indicates areas with no data where communal farmlands or plantation forests were removed.}\label{afoto}
\end{figure}

The most pronounced errors were in the MODIS and GlobCover maps, which both underestimated cropland extent by 10-75\% in the center of the country (blue areas in Fig. 1, and the dominant production region), and overestimated along the eastern to northern margins (red areas in Fig. 1). Averaging these errors across the whole country shows resulted

The average of these were 21\% for MODIS and 34\% for GlobCover at 1 km resolution (Fig. S1), meaning that each map had a strong underestimation bias. Both maps's biases decreased with pixel aggregation, with MODIS biases falling to 8\% at 50 km, whereas GlobCover's was still 24\% at 100 km (SI Appendix, Fig. S1).

The SA-LC map uniformly overestimated cropland throughout the country (Fig. 1), but its overall bias was relatively small, ranging from -8\% at 1 km to -6\% at 100 km  (Fig. S1). The GeoWiki map had a strongly heterogenous pattern of differences (Fig. 1) and the smallest magnitude of bias, which changed between slight tendencies to underestimate (5\% at 1 km) and overestimate (-2 \% at 100 km,  Fig. S1).

\subsection{Landscape characteristics and error magnitude}
To investigate the degree to which landscape features influence landcover map errors, we extracted all pixels in agricultural areas ($>$0\% cropland) of the 1 km reference map using the boundaries of 354 magisterial districts (South Africa's finest administrative unit, which average 3,445 km$^2$ in size; SI Appendix, Fig. S3), and calculated the district-wise mean for these pixels, providing a measure of cropland density that was informative of the degree of mixing between cropland and other land covers at landscape scales. We then extracted the cropland map errors for the same pixels, and calculated their mean absolute errors for each district.

A generalized additive model fit to district-level mean absolute error (log-transformed) shows that error peaks at 50\% cropland cover for all but the GlobCover map (which continued to increase with cropland cover), and is lowest when the landscape is dominated either by cropland or other cover types (Fig. 2). In other words, bias is highest when cropland cover is mixed evenly with other cover types. The reason that GlobCover bias continued to increase with cropland density was because this dataset's dominant class identified for South African cropland was a mixture of 50-70\% cropland and other vegetation types, which resulted in the capping of GlobCover-derived cropland cover estimates at the intermediate densities defined by this class.

%\vspace{-1 cm}
\begin{figure}[ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
\caption{The relationship between the mean absolute error (error magnitude) in cropland maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model. Prediction curves are color-coded to the different cropland maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}
\label{afoto}
\end{figure}

\subsection{Map error and Tier-1 carbon estimates}
Using the methods provided by \cite{ruesch_new_2008}, we calculated average carbon densities for African forests, secondary forests, shrublands, croplands, grasslands, and sparse habitats (semi-arid grasslands and low shrublands), and assigned cropland carbon values to map cells in proportion to their cropland cover. For the non-cropland proportions, we assigned the carbon value from each of the other types, creating five different carbon maps for each landcover map at each aggregation scale (Fig. S4), which allowed us to test how carbon estimates vary as a function of i) cropland map bias and ii) the characteristics of adjacent cover types.

@@

The spatial patterns of errors in carbon estimates (Fig. S4) reflect those of cropland biases (Fig. 1). Where cropland was underestimated and the surrounding cover type was of higher carbon density than cropland, carbon density was overestimated. For lower density cover (grassland and sparse vegetation), carbon stocks were underestimated, but by small magnitudes. These tendencies were reflected in each map's biases, as calculated over the cropped areas of the country as jointly defined by the reference and each cropland map (Fig. 3). For example, MODIS and GlobCover bias was $\sim$-50\% (overestimation) at 1 km resolution when forest was the cropland-adjacent cover (stars in semi-transparent green and gold bars, Fig. 3; Table S2). For sparse vegetation (open circles in Fig. 3), MODIS bias was 3\% at all scales, whereas maps that overestimated cropland (e.g. SA-LC, semi-transparent red) overestimated carbon density for this cover type, because cropland has a higher carbon density \cite{ruesch_new_2008}. Overall, GeoWiki had the lowest bias, for all cover types and all resolutions. Its worst bias was a tendency to overestimate by 12\% at 1 km when forest was adjacent, but at coarser scales this bias reduced to just a few percent (Fig. 3, Table S2). All maps' biases are within +/-10\% bias after aggregation to 25 km.

%\vspace{-0.9 cm}
\begin{figure}[!ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scale.pdf}}
\caption{Biases and mean absolute errors in carbon densities derived from cropland maps, calculated as percents relative to the reference map. Bias estimates (represented by symbols) fall within the semi-transparent bars, while mean absolute errors are contained in the solid bars. Bar colors are coded to specific cropland map, while the symbols indicate which cover type was used to calculate cropland adjacent carbon density. The bar represents the mean biases calculated across each of the 5 cover types. Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).}
\label{afoto}
\end{figure}

The mean absolute error in carbon maps (solid colored bars in Fig. 3) generally followed the same patterns, but with higher magnitudes and a few important differences. The most notable is that Geo-Wiki, despite relatively low bias, had large mean absolute errors at 1 km, averaging 27\% across across cover types (line in solid blue bar, Fig. 3, Table S2), which is close to the 36-37\% for GlobCover and MODIS. SA-LC had the lowest mean absolute error across scales, averaging (across cover types) 14\% at 1 km to 3\% at 100 km. (Fig. 3, Table S2). The increase in Geo-Wiki's absolute bias relative to SA-LC's can be attributed to the highly heterogeneous nature of its cropland errors (Fig. 1), which traded between between positive and negative errors of high magnitude over shorter distances than the other three cropland maps.

\subsection{Map error and gridded agricultural data}
The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived cropland area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}. To be consistent with this methodology, we first adjusted our cropland maps according to this procedure, using the reference map to calculate total cropland area for each of South Africa's nine provinces, then updating the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}. Despite this statistical constraint, the updated cropland maps still had substantial errors that were similar in pattern (Fig. S5) to those in the unadjusted maps (Fig. 1), and we evaluated how these residuals affected gridded estimates of the yield and production of maize, South Africa's largest crop \cite{estes_comparing_2013}. To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize (South Africa's largest crop by area, \cite{estes_comparing_2013}) yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.

@@

%\vspace{-1 cm}
\begin{figure}[!ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/yield_bias_scale.pdf}}
\caption{Biases and mean absolute errors in disaggregated maize yield and production estimates. Bias estimates (represented by symbols) fall within the semi-transparent bars, while mean absolute errors are contained in the solid bars. Bar colors are coded to specific cropland map, while the symbols indicate which the values for the different variables (production and yield).}
\label{afoto}
\end{figure}

@@

\subsection{Map error and evapotranspiration}
Compared to carbon and crop related examples, bias and mean absolute errors in evapotranspiration (ET) calculated using the VIC model were small and averaged to less than +/-1\%. However, there were several hotspots of discrepancy evident in the error maps (Fig. 6). The most pronounced of these are the 5-15\% overestimates in the center resulting from VIC when initialized with MODIS and GlobCover, while overestimates along the southern and western coasts reached 25\%. These locations correspond primarily to the margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, marked approximately by the 400 mm isohyet, where maize is the primary crop. The west coast hotspot falls at the western edge of the wheat-dominated winter rainfall region \cite{hardy_rainfed_2011}, where growing season rainfall is approximately 200 mm.

@@

\subsection{Initialization errors in agent-based models}
We used an agent-based model (ABM) of food security that represents the interactions between hundreds of individual farming households over multiple seasons \cite{chen_dependency_2013}. Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g. districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) that are needed to represent the different land units of single farmers. To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig. S8) in the eastern part of the country, having between 28-45\% of their areas devoted to cropland, according to the reference map.

We disaggregated the cropland percentages in all maps to binary cropland/non-cropland cover types with 100 m resolution, which matches the typical field size (1 ha) for smallholder farmers in household survey data (collected in Zambia) used in developing the agent-based model \cite{chen_dependency_2013}. These surveys found mean household crop field area to be 2 ha, which we divided into reference cropland areas to estimate the total households within each district. We then initialized the model by assigning each household agent two cropland pixels. In order to emulate the natural groupings of communities, the model only assigns a household fields that are within 1.5 km of other agents' fields, provided those pixels were not previously allocated to another agent. The model thus iteratively grows ``communities'' until all households are assigned cropland, or all available cropland is allocated.

>>>>

%\usepackage{pnastwoF}
%\usepackage{cite}
\usepackage[numbers,round,sort,compress]{natbib}
\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
\usepackage{lineno}
\setlength\linenumbersep{2pt}
\usepackage{placeins}
\usepackage{float}
%\usepackage{subfigure}
\usepackage{wrapfig}
%\setlength\belowcaptionskip{-2cm}
%\setlength{\textfloatsep}{5pt plus 1.0pt minus 2.0pt}

@@

Being unable to fully quantify the errors in landcover maps of course makes it difficult, if not impossible, to quantify their impact on downstream analyses. There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, and carbon flux \cite{quaife_impact_2008} and human population estimates \cite{linard_assessing_2010}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}. One exception is a Belgian study  \cite{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover map errors bias measurements of agricultural land use patterns, but the study extent was fairly small and the validation data were discontiguous.

Just as a building needs a solid foundation, global change science needs to be based on sound landcover data. There is thus an urgent need to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.  We address this need here, using a unique, high accuracy agricultural landcover map for the entire country of South Africa to conduct a spatially comprehensive, bottom-up quantification of error in several latest generation landcover maps that are widely used in global change studies. We use these errors to assess the biases and accuracies of i) these landcover data, ii) how landcover properties influence map errors, iii) how bias and accuracy change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and iv) how landcover error propagates through ``downstream'' analyses that represent several major global change research areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.

@@

The South African government commissioned a whole-country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}. The map was made by trained workers who visually interpreted high resolution satellite imagery ($<$5 m SPOT imagery) and manually digitized field boundaries following a standardized mapping protocol. The resulting vectorized field maps, provide a unique, high quality reference dataset describing South African crop field distributions and size classes for the period 2009-2011, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution. We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create a gridded cropland reference map.

We compared the reference map with similar maps derived from four existing landcover datasets. We obtained South Africa's 30 m resolution National Landcover map (SA-LC) for 2009 \cite{sanbi_national_2009}, the 500 m resolution MODIS Landcover for 2011 \cite{land_processes_distributed_active_archive_center_lp_daac_modis_2011, friedl_modis_2010}, the 300 m resolution GlobCover 2009 \cite{arino_global_2012}, and the new 1 km Geo-wiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}. We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite[e.g.][]{fry_completion_2009}, MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013, shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and represents the current state-of-the-art in landcover mapping. We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates, resulting in 4 maps (hereafter simply the ``cropland maps'') to compare to our reference map.

\vspace{-0.5 cm}
\section{Quantifying Error and Bias}
We first calculated errors in pixel-wise estimated cropland percentages by differencing the reference map with each cropland map at each of the five aggregation scales (1-100 km). We used these errors to calculate map bias (the mean pixel-wise error) and accuracy (the mean absolute error, and how these vary with scale. Next, we assessed the degree to which the average cropland cover in agricultural landscapes, a descriptor of landcover pattern, impacts map accuracy.

To investigate how landcover map error can impact downstream global change research, we quantified the biases and accuracies in five landcover-based analyses built upon our reference and cropland maps. The first three represented ``first-order'' analyses, in which a variable of interest is mapped onto a landcover type(s) using a simple empirical relationship. The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}. The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates. Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.

The other two analyses can be considered second-order, wherein a process model draws on the cover types' values to calculate an output value. For the first of these, we used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model to calculate monthly evapotranspiration, using the reference and cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components. In the second example, we examined how these map errors impact the land allocation process of an agent-based food security model \cite{chen_dependency_2013}.

\vspace{-0.5 cm}
\section{Landcover maps}
\subsection{Bias and accuracy}

We created the 1 km reference map after removing all field types classified as communal/smallholder agriculture (individual fields in this category were not delineated, thus they were removed to prevent potential cropland area overestimates, SI) or permanent tree-crops (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa). The 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.

We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from each reference (2007 and 2011) map at each scale of aggregation to assess error patterns (Fig. 1A). Negative pixels here represent overestimation error, while positive values indicate underestimates.

%\vspace{-0.5 cm}
\begin{figure*}[ht]
%\begin{wrapfigure}{c}{1\textwidth}
\centerline{\includegraphics[width=0.99\textwidth]{figures/figure1.pdf}}
\caption{(A) Errors in the percent cropland estimates resulting from each of the four cropland maps relative to the reference map at different scale of pixel aggegation. Rows indicate the landcover map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation. White indicates areas with no data where communal farmlands or plantation forests were removed from analysis. (B) The bias (mean error) and accuracy (mean absolute error [MAE]) of each cropland map at each scale of aggregation. Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps. The symbols indicate the method of calculating each metric, either by averaging across the entire country, within the union of  agricultural areas (cropland $>$0.05\%) for each reference-cropland map combination, or as the average of the averages calculated per 5\% bin of cropland cover, a measure of error that is independent of cropland density.}\label{afoto}
\end{figure*}
%\end{wrapfigure}

%\begin{figure*}[htp]
%\centering
%    \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}\quad
%     \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}
%\end{figure*}

\linenumbers
The most pronounced errors were in the MODIS and GlobCover maps, which both underestimated cropland extent by 10-75\% in the center of the country (blue areas in Fig. 1A, and the dominant production region), and overestimated along the eastern to northern margins (red areas in Fig. 1A). All maps have less than +/-5\% bias (crosses in semi-transparent bars, Fig. 1B), at all aggregation scales, when errors are averaged across the entire country, due to the extensive non-agricultural areas in the country that were effectively discerned by all maps. The whole country mean absolute error (the measure of map accuracy) were a few percent higher values in most cases, but still $<$10\%, except for GlobCover ($\sim$11\%) at 1 km resolution. Biases increase only slightly when errors are averaged within the combined agricultural regions (areas with greater than 0.05\% cropland, indicated by circles in transparent bars in Fig. 1B) of the reference and each cropland map, but accuracy decreases substantially, with mean absolute errors increasing to 10\% for SA-LC and around 20\% for MODIS, GlobCover, and GeoWiki at 1 km resolution (circles in solid bars in Fig. 1B), and remain between 5-10\% for all three datasets even at 25 km of aggregation.

As a third measure, we calculated bias and mean absolute error by averaging their mean per-ventile values (i.e. the average bias or accuracy value for areas have 0-5\% cropland, 5-10\% cropland, etc.). These provide density independent measures of map performance because they remove the dominance any particular level of cover has on map accuracy, which in this case was the 0-5\% cover class where map performance was best (SI). Calculated this way, bias and accuracy were substantially worse. The average bias jumped to 21\% for MODIS and 34\% for GlobCover at 1 km resolution (Fig. 1B), meaning that each map had a strong tendency to underestimate cropland. This decreased with pixel aggregation, falling to 8\% at 50 km for MODIS, but remaining as high as 14\% at 100 km for GlobCover. SA-LC bias was smaller in magnitude, but tended towards cropland overestimates of 5-10\% at all aggregation scales. GeoWiki alone was relatively unbiased, but showed substantial inaccuracy with a density-independent mean absolute error of 23\%, which dropped below 10\% only above 25 km of aggregation. GlobCover was by far the most inaccurate, having $>$35\% mean absolute error at 1 km and 17\% at 100 km, following by MODIS (over 30\% at 1 km and 11\% at 100 km). SA-LC was the most accurate of all four maps with the density-independent measure, ranging from just above 10\% mean absolute error at 1 km to 5\% at 100 km.

\subsection{Landscape characteristics and error}
The density-independent measures indicate that error and bias increase with cropland density, which suggests that map error varies as a function of the typical abundance of cover in the landscape being imaged. To investigate the relationship between cover density and map accuracy, we calculated the mean density for agricultural pixels ($>$0.05\% cropland) in the 1 km reference map within the boundaries of 354 magisterial districts (South Africa's finest administrative unit, averaging 3,445 km$^2$; SI Figure S3), providing a landscape-scale measure representing typical cropland density. We then extracted the cropland map errors for these same pixels, and calculated their district-wise mean absolute errors.

A generalized additive model fit to district-level mean absolute error (log-transformed) shows that error peaks at 50-60\% cropland cover for all but the GlobCover map (which continued to increase with cropland cover), and is lowest when the landscape is dominated either by cropland or other cover types (Fig. 2). In other words, accuracy is generally lowest when cropland cover is mixed evenly with other cover types. GlobCover's accuracy continued to decrease with cropland density because the dominant cover class contributing to the percentage cropland estimate was a mixed class defined as 50-70\% crops mingled with other vegetation, thus the maximum percentage was constrained by this mixture range.

%\vspace{-1 cm}
\begin{figure}[ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
\caption{The relationship between map accuracy (the mean absolute error) in cropland maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model. Prediction curves are color-coded to the different cropland maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}
\label{afoto}
\end{figure}

\section{First-order analyses}
\subsection{Tier-1 carbon estimates}
Using the methods provided by \cite{ruesch_new_2008}, we calculated average carbon densities for African forests, secondary forests, shrublands, croplands, grasslands, and sparse habitats (semi-arid grasslands and low shrublands), and assigned cropland carbon values to map cells in proportion to their cropland cover. For the non-cropland proportions, we assigned the carbon value from each of the other types, creating five different carbon maps for each landcover map at each aggregation scale (Fig. S4), which allowed us to test how carbon estimates vary as a function of i) cropland map bias and ii) the characteristics of adjacent cover types.

@@

The spatial patterns of errors in carbon estimates (Fig. S4) reflect those of the cropland maps (Fig. 1). Where cropland was underestimated and the surrounding cover type was of higher carbon density than cropland, carbon density was overestimated. For lower density cover (grassland and sparse vegetation), carbon stocks were underestimated, but by small magnitudes. These tendencies were reflected in each map's biases, as calculated over the cropped areas of the country as jointly defined by the reference and each cropland map (Fig. 3). For example, MODIS and GlobCover bias was $\sim$-50\% (overestimation) at 1 km resolution when forest was the cropland-adjacent cover (stars in semi-transparent green and gold bars, Fig. 3; Table S2). For sparse vegetation (open circles in Fig. 3), MODIS bias was 3\% at all scales, whereas maps that overestimated cropland (e.g. SA-LC, semi-transparent red) overestimated carbon density for this cover type, because cropland has a higher carbon density \cite{ruesch_new_2008}. Overall, GeoWiki had the lowest bias, for all cover types and all resolutions. Its worst bias was a tendency to overestimate by 12\% at 1 km when forest was adjacent, but at coarser scales this bias reduced to just a few percent (Fig. 3, Table S2). All maps' biases are within +/-10\% bias after aggregation to 25 km.

%%\vspace{-0.9 cm}
%\begin{figure}[!ht]
%\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scale.pdf}}
%\caption{Biases and mean absolute errors in carbon densities derived from cropland maps, calculated as percents relative to the reference map. Bias estimates (represented by symbols) fall within the semi-transparent bars, while mean absolute errors are contained in the solid bars. Bar colors are coded to specific cropland map, while the symbols indicate which cover type was used to calculate cropland adjacent carbon density. The bar represents the mean biases calculated across each of the 5 cover types. Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).}
%\label{afoto}
%\end{figure}

The mean absolute error in carbon maps (solid colored bars in Fig. 3) generally followed the same patterns, but with higher magnitudes and a few important differences. The most notable is that GeoWiki, despite relatively low bias, was highly inaccurate at 1 km, with 27\% mean absolute error across across cover types (line in solid blue bar, Fig. 3, Table S2), which is close to the 36-37\% for GlobCover and MODIS. SA-LC had the lowest mean absolute error across scales, averaging (across cover types) 14\% at 1 km to 3\% at 100 km. (Fig. 3, Table S2). The decrease in GeoWiki's accuracy relative to SA-LC's can be attributed to the highly heterogeneous nature of its cropland errors (Fig. 1), which alternated between high magnitude positive and negative errors over relative short distances.

\subsection{Gridded yield and production estimates}
The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived cropland area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}. To be consistent with this methodology, we first adjusted our cropland maps according to this procedure, using the reference map to calculate total cropland area for each of South Africa's nine provinces, then updating the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}. Despite this statistical constraint, the updated cropland maps still had substantial errors that were similar in pattern (Fig. S5) to those in the unadjusted maps (Fig. 1), and we evaluated how these residuals affected gridded estimates of the yield and production of maize, South Africa's largest crop \cite{estes_comparing_2013}. To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize (South Africa's largest crop by area, \cite{estes_comparing_2013}) yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.

@@

%\vspace{-1 cm}
\begin{figure*}[ht]
\centerline{\includegraphics[width=1\textwidth]{figures/figure3.pdf}}
\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in (A) carbon densities derived from cropland maps and (B) in disaggregated maize yield and production estimates. Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps. For the carbon maps (A), symbols indicate the cover type used to calculate cropland-adjacent carbon density: the line represents the mean bias/MAE across all 5 cover types; shrubland and grassland values (near zero) and secondary forest (close to forest values) are not shown for clarity (see SI). For (B) symbols code the different variables (production and yield), normalized to their respective means. Values in A and B were calculated within the union of agricultural areas (cropland $>$0.05\%) for each reference-cropland map combination.}
\label{afoto}
\end{figure*}

@@

\section{Second-order analyses}
\subsection{Evapotranspiration estimates}
Compared to carbon and crop related examples, bias and accuracy in evapotranspiration (ET) calculated using the VIC model were small and averaged to less than +/-1\%. However, there were several error hotspots in the resulting ET difference maps (Fig. 6). The most pronounced of these are the 5-15\% overestimates in the center resulting from VIC when initialized with MODIS and GlobCover, while overestimates along the southern and western coasts reached 25\%. These locations correspond primarily to the margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, marked approximately by the 400 mm isohyet, where maize is the primary crop. The west coast hotspot falls at the western edge of the wheat-dominated winter rainfall region \cite{hardy_rainfed_2011}, where growing season rainfall is approximately 200 mm.

@@

\begin{figure}[!ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.}\label{afoto}
\end{figure}
